---
title: "ML Pima Indians Diabetes"
author: "Michael Harrison"
date: "May 2, 2017"
output: html_document
---

# Load Libraries and data
```{r}
library(caret)
library(mlbench)
library(corrplot)
library(e1071)
library(parallel)
library(doParallel)
data("PimaIndiansDiabetes")
```
#Dataset dimensions
```{r}
dim(PimaIndiansDiabetes)
```
#Partition dataset 
```{r}
inTrain <- createDataPartition(PimaIndiansDiabetes$diabetes, p=0.80, list=FALSE)
training <- PimaIndiansDiabetes[inTrain,]
testing <- PimaIndiansDiabetes[-inTrain,]
dim(training)
```
```{r}
sapply(training, class)
```
```{r}
head(training)
```
```{r}
summary(training)
```
#Class distribution
```{r}
cbind(freq = table(training$diabetes), percentage = prop.table(table(training$diabetes)))
```
# Checking for attribute correlation
```{r}
cor(training[,1:8])
```

#Data Visualiztion
- Attribute histogram
```{r}
par(mfrow = c(2,4))
for(i in 1:8){
        hist(training[,i], main = names(training)[i], xlab = names(training)[i])
}
```
# Attribute Density
```{r}
par(mfrow = c(2,4))
for(i in 1:8){
        plot(density(training[,i]), main = names(training)[i], xlab = names(training)[i])
}
```
# Skew check
```{r}
apply(training[,1:8], 2, skewness)
```

#Box and Whisker
```{r}
par(mfrow = c(2,4))
for(i in 1:8){
        boxplot(training[,i], main = names(training)[i], xlab = names(training)[i])
}
```

#Correlation Plot
```{r}
correlations <- cor(training[,1:8])
corrplot(correlations)
```

```{r}
pairs(diabetes~., data=training, col=training$diabetes)
```
#Class wise density plot
```{r}
scales <- list(x = list(relation = "free"), y = list(relation = "free"))
featurePlot(x = training[,1:8], y = training[,9],
            plot="density", scales = scales)
```
#Box and Whiskey by Class
```{r}
featurePlot(x = training[,1:8], y = training[,9], 
            plot = "box", col = training$diabetes)
```

#Training Harness
```{r}
seed <- set.seed(13)
fitControl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
metric <- "Accuracy"
```


#Baseline Algorithm Suite
```{r}
#General Linearized Model
set.seed(seed)
fitGLM <- train(diabetes~., data = training, 
             method = "glm", metric = metric,
             trControl = fitControl)
#General Linearized Model 
set.seed(seed)
fitGLMNET <- train(diabetes~., data = training, 
             method = "glmnet", metric = metric,
             trControl = fitControl)
#Bayesian Generalized Linear Model
set.seed(seed)
fitBGLM <- train(diabetes~., data = training, 
             method = "bayesglm", metric = metric,
             trControl = fitControl)
#Support Vector Machines w/ Radial Basis Function
set.seed(seed)
fitSVM <- train(diabetes~., data = training, 
             method = "svmRadial", metric = metric,
             trControl = fitControl)
#Support Vector Machines w/ Class Weights
set.seed(seed)
fitCWSVM <- train(diabetes~., data = training, 
             method = "svmRadialWeights", metric = metric,
             trControl = fitControl)
#Least Squares Support Vector Machines with Radial Basis Function
set.seed(seed)
fitLSSVM <- train(diabetes~., data = training, 
             method = "lssvmRadial", metric = metric,
             trControl = fitControl)
#Naive Bayes
set.seed(seed)
fitNB <- train(diabetes~., data = training, 
             method = "nb", metric = metric,
             trControl = fitControl)

results <- resamples(list(GLM = fitGLM,
                          GLMNET = fitGLMNET,
                          BGLM = fitBGLM,
                          SVM = fitSVM,
                          CWSVM = fitCWSVM,
                          LSSVM = fitLSSVM,
                          NB = fitNB))

dotplot(results, scales = scales, main = "Baseline")
```
```{r}
summary(results)
```

# Algorithm Suite w/ Preprocessing: BoxCox
```{r}
#Engage parallel processing
cluster <- makeCluster(detectCores()-1)
registerDoParallel(cluster)

#General Linearized Model
set.seed(seed)
fitGLM <- train(diabetes~., data = training, 
             method = "glm", metric = metric,
             preProc = c("BoxCox"),
             trControl = fitControl)
#General Linearized Model 
set.seed(seed)
fitGLMNET <- train(diabetes~., data = training, 
             method = "glmnet", metric = metric,
             preProc = c("BoxCox"),
             trControl = fitControl)
#Bayesian Generalized Linear Model
set.seed(seed)
fitBGLM <- train(diabetes~., data = training, 
             method = "bayesglm", metric = metric,
             preProc = c("BoxCox"),
             trControl = fitControl)
#Support Vector Machines w/ Radial Basis Function
set.seed(seed)
fitSVM <- train(diabetes~., data = training, 
             method = "svmRadial", metric = metric,
             preProc = c("BoxCox"),
             trControl = fitControl)
#Support Vector Machines w/ Class Weights
set.seed(seed)
fitCWSVM <- train(diabetes~., data = training, 
             method = "svmRadialWeights", metric = metric,
             preProc = c("BoxCox"),
             trControl = fitControl)
#Least Squares Support Vector Machines with Radial Basis Function
set.seed(seed)
fitLSSVM <- train(diabetes~., data = training, 
             method = "lssvmRadial", metric = metric,
             preProc = c("BoxCox"),
             trControl = fitControl)
#Naive Bayes
set.seed(seed)
fitNB <- train(diabetes~., data = training, 
             method = "nb", metric = metric,
             preProc = c("BoxCox"),
             trControl = fitControl)

#Disengage parallel processing
stopCluster(cluster)
registerDoSEQ()

transformResults <- resamples(list(GLM = fitGLM,
                          GLMNET = fitGLMNET,
                          BGLM = fitBGLM,
                          SVM = fitSVM,
                          CWSVM = fitCWSVM,
                          LSSVM = fitLSSVM,
                          NB = fitNB))

dotplot(transformResults, scales = scales, main = "Transform1: BoxCox")
```
```{r}
summary(transformResults)
```

#Algorithm Suite w/ Preprocessing: center, scale, BoxCox
```{r}
#Engage parallel processing
cluster <- makeCluster(detectCores()-1)
registerDoParallel(cluster)

#General Linearized Model
set.seed(seed)
fitGLM <- train(diabetes~., data = training, 
             method = "glm", metric = metric,
             preProc = c("center", "scale", "BoxCox"),
             trControl = fitControl)
#General Linearized Model 
set.seed(seed)
fitGLMNET <- train(diabetes~., data = training, 
             method = "glmnet", metric = metric,
             preProc = c("center", "scale", "BoxCox"),
             trControl = fitControl)
#Bayesian Generalized Linear Model
set.seed(seed)
fitBGLM <- train(diabetes~., data = training, 
             method = "bayesglm", metric = metric,
             preProc = c("center", "scale", "BoxCox"),
             trControl = fitControl)
#Support Vector Machines w/ Radial Basis Function
set.seed(seed)
fitSVM <- train(diabetes~., data = training, 
             method = "svmRadial", metric = metric,
             preProc = c("center", "scale", "BoxCox"),
             trControl = fitControl)
#Support Vector Machines w/ Class Weights
set.seed(seed)
fitCWSVM <- train(diabetes~., data = training, 
             method = "svmRadialWeights", metric = metric,
             preProc = c("center", "scale", "BoxCox"),
             trControl = fitControl)
#Least Squares Support Vector Machines with Radial Basis Function
set.seed(seed)
fitLSSVM <- train(diabetes~., data = training, 
             method = "lssvmRadial", metric = metric,
             preProc = c("center", "scale", "BoxCox"),
             trControl = fitControl)
#Naive Bayes
set.seed(seed)
fitNB <- train(diabetes~., data = training, 
             method = "nb", metric = metric,
             preProc = c("center", "scale", "BoxCox"),
             trControl = fitControl)

#Disengage parallel processing
stopCluster(cluster)
registerDoSEQ()

transform2Results <- resamples(list(GLM = fitGLM,
                          GLMNET = fitGLMNET,
                          BGLM = fitBGLM,
                          SVM = fitSVM,
                          CWSVM = fitCWSVM,
                          LSSVM = fitLSSVM,
                          NB = fitNB))

dotplot(transform2Results, scales = scales, main = "Transform2: Center, Scale, BoxCox")
```

```{r}
summary(transform2Results)
```

# Algorithm Suite w/ Preproceesing: center, scale
```{r}
#Engage parallel processing
cluster <- makeCluster(detectCores()-1)
registerDoParallel(cluster)

#General Linearized Model
set.seed(seed)
fitGLM <- train(diabetes~., data = training, 
             method = "glm", metric = metric,
             preProc = c("center", "scale"),
             trControl = fitControl)
#General Linearized Model 
set.seed(seed)
fitGLMNET <- train(diabetes~., data = training, 
             method = "glmnet", metric = metric,
             preProc = c("center", "scale"),
             trControl = fitControl)
#Bayesian Generalized Linear Model
set.seed(seed)
fitBGLM <- train(diabetes~., data = training, 
             method = "bayesglm", metric = metric,
             preProc = c("center", "scale"),
             trControl = fitControl)
#Support Vector Machines w/ Radial Basis Function
set.seed(seed)
fitSVM <- train(diabetes~., data = training, 
             method = "svmRadial", metric = metric,
             preProc = c("center", "scale"),
             trControl = fitControl)
#Support Vector Machines w/ Class Weights
set.seed(seed)
fitCWSVM <- train(diabetes~., data = training, 
             method = "svmRadialWeights", metric = metric,
             preProc = c("center", "scale"),
             trControl = fitControl)
#Least Squares Support Vector Machines with Radial Basis Function
set.seed(seed)
fitLSSVM <- train(diabetes~., data = training, 
             method = "lssvmRadial", metric = metric,
             preProc = c("center", "scale"),
             trControl = fitControl)
#Naive Bayes
set.seed(seed)
fitNB <- train(diabetes~., data = training, 
             method = "nb", metric = metric,
             preProc = c("center", "scale"),
             trControl = fitControl)

#Disengage parallel processing
stopCluster(cluster)
registerDoSEQ()

transform3Results <- resamples(list(GLM = fitGLM,
                          GLMNET = fitGLMNET,
                          BGLM = fitBGLM,
                          SVM = fitSVM,
                          CWSVM = fitCWSVM,
                          LSSVM = fitLSSVM,
                          NB = fitNB))

dotplot(transform3Results, scales = scales, main = "Transform3: Center, Scale")
```
```{r}
summary(transform3Results)
```


#Algorithm Tuning
-Tuning Harness
```{r}
fitControl <- trainControl(method = "repeatedcv", number = 10, repeats =3)
metric <- "Accuracy"
```


-GLMNET, no transform 
```{r}
#General Linearized Model 
set.seed(seed)
fitGLMNET0 <- train(diabetes~., data = training, 
             method = "glmnet", metric = metric,
             trControl = fitControl)
fitGLMNET0
```
```{r}
grid <- expand.grid(.alpha = seq(.4, 1, by = .05),
                    .lambda = c(0.0004548649, 0.0045486487, 0.0454864867))
set.seed(seed)
tuneGLMNET <- train(diabetes~., data = training, 
             method = "glmnet", metric = metric,
             tuneGrid = grid,
             trControl = fitControl)
plot(tuneGLMNET)
```
```{r}
tuneGLMNET$bestTune
```


-CWSVM, transform1 (center, scale, BoxCox)
```{r}
set.seed(seed)
fitCWSVM1 <- train(diabetes~., data = training, 
             method = "svmRadialWeights", metric = metric,
             preProc = c("center", "scale", "BoxCox"),
             trControl = fitControl)
fitCWSVM1
```
```{r}
grid <- expand.grid(.sigma = 0.1155233,
                    .C = seq(.30, .70, by = .05),
                    .Weight = c(1,2,3))
set.seed(seed)
tuneCWSVM <- train(diabetes~., data = training, 
             method = "svmRadialWeights", metric = metric,
             tuneGrid = grid,
             preProc = c("center", "scale", "BoxCox"),
             trControl = fitControl)
plot(tuneCWSVM)
```
```{r}
tuneCWSVM
```



-LSSVM, no transform
```{r}
set.seed(seed)
fitLSSVM0 <- train(diabetes~., data = training, 
             method = "lssvmRadial", metric = metric,
             trControl = fitControl)
fitLSSVM0
```

```{r}
grid <- expand.grid(.sigma = seq(.03, .3, by = .1),
                    .tau = seq(.1, .24, by = .02))
set.seed(seed)
tuneLSSVM <- train(diabetes~., data = training, 
             method = "lssvmRadial", metric = metric,
             tuneGrid = grid,
             trControl = fitControl)
plot(tuneLSSVM)
```
```{r}
tuneLSSVM
```

